---
title: "Upload and Download Formats: Notes on Round-trips"
date: "December 24, 2016"
output: html_document
---

Google docs allows documents to be uploaded and downloaded in various formats, and happily, many of these overlap usefully with the formats offered by pandoc.

However, not all features of a document survive the 'round trip', being converted from markdown, to the upload format, to a Google doc, to the download format, and then back to markdown.

This document goes through each combination of upload and download format[^1], and runs some simple tests to determine if the integrity of the underlying document is maintained.

```{r}
library(tidyverse)
library(brocks)
library(googdown)

get_ast <- function(x) {
  system(paste0("pandoc ", x, " -t json"), intern = TRUE)
}

# A function to extract the (unique) names of the AST elements contained in a
# document
ast_elements <- function(json_string) {
  # Read in the json string as a list, and then flatten it
  elements <- unlist(jsonlite::fromJSON(json_string, simplifyVector = FALSE))
  
  # Find the unique content types, and return them alphabetical order
  sort(unique(elements[grepl(".*t", names(elements))]))
}

# A function to return a bool; do two documents have *extactly the same* AST?
same_ast <- function(x, y) {
  x_ast <- get_ast(x)
  y_ast <- get_ast(y)
  
  x_ast == y_ast
}

# A function to return a bool; Are all AST elements in X, also in Y?
same_ast_elements <- function(x, y) {
  x_ast <- get_ast(x)
  y_ast <- get_ast(y)
  
  all(ast_elements(x_ast) %in% ast_elements(y_ast))
}
```

```{r}
# Some code to try out a few different features with different combinations of
# readers and writers

setwd("consistency_tests")

# Init a data.frame to store results in
results <- na.omit(data.frame(
  test              = NA, 
  upload_format     = NA, 
  download_format   = NA, 
  same_ast          = NA,
  same_ast_elements = NA
))

# Get vectors of the parameters which we'd like to try-out together
upload_formats   <- names(googdown:::file_types())
download_formats <- googdown:::reader_list()
test_files       <- normalizePath(list.files(
  "test_cases", full.names = TRUE
))


setwd("results")

# These tests are time-consuming. If the tests have been run in the same format
# before, fall-back to loading a cached-file of the results. The 'same tests' is
# defined as the same upload and download formats, and the same test cases.

input_hash <- digest::digest(
  c(lapply(test_files, function(x) openssl::md5(read_txt(x))),
    download_formats,
    upload_formats),
  algo = "md5"
)

d <- dir.create(".cached_results")
cache_dir  <- normalizePath(".cached_results")
cache_file <- paste0(cache_dir, "/", input_hash, ".rds")

gd_auth()

# Obtain results
if (file.exists(cache_file)) {
  # If you've already run these tests, read-in the results from last time
  results <- readRDS(cache_file)
} else {
  # Iterate through the combinations of upload and download formats, checking if
  # the AST (and AST elements) have been maintained
  for (u in upload_formats) {
    for (d in download_formats) {
      # Create a dir to store the results in
      result_dir <- paste0(u, "-", d)
      dir.create(result_dir)
      setwd(normalizePath(result_dir))
      
      for (t in test_files) {
        # Upload the test file to google docs
        upload <- gd_upload(file_name = t, format = u)
        
        # Download the test file
        returned_file_name <- paste0("returned-", basename(t))
        download <- gd_download(
          file_id = upload$id, file_name = returned_file_name,
          format = d, output_format = "markdown"
        )
        
        # See if it came out the same, put the results in the output data.frame
        results <- bind_rows(
          results, 
          data.frame(
            test              = basename(t),
            upload_format     = u, 
            download_format   = d,
            same_ast          = same_ast(t, returned_file_name),
            same_ast_elements = same_ast_elements(t, returned_file_name)
          )
        )
        
        # Delete the google doc, to keep the account managable
        gd_delete(upload$id)
      }
      # Pop a level up the dir structure
      setwd("..")
    }
  }
  # Save the results for next time
  saveRDS(results, cache_file)
} 

# Print out the whole darn data.frame
results <- results %>% 
  mutate(
    combo_id = as.numeric(factor(paste0(upload_format, download_format)))
  )

```


## Features retained (or not)

```{r}
results %>% 
  group_by(test) %>% summarise(mu = mean(same_ast)) %>%
  arrange(desc(mu))

results %>% 
  group_by(test) %>% summarise(mu = mean(same_ast_elements)) %>%
  arrange(desc(mu))

```

Features with no equivalent in Google docs:

- Code, in-line
- Code, in-blocks
- Figure captions
- Table captions

### Basic text (pargraphs and formatting)

Regardless of input format, using HTML as an output format loses bold, italic, super-script and sub-script.

```{r}
results %>% filter(test == "bold_italic.md")

read_txt("../test_cases/hyperlink.md")
read_txt("open_office_doc-ms_word_doc/returned-hyperlink.md")
read_txt("ms_word_doc-open_office_doc/returned-hyperlink.md")


cat(read_txt("../test_cases/list_bulleted.md"))
cat(read_txt("open_office_doc-ms_word_doc/returned-list_bulleted.md"))
cat(read_txt("ms_word_doc-open_office_doc/returned-list_bulleted.md"))
cat(read_txt("ms_word_doc-ms_word_doc/returned-list_bulleted.md"))

```




### Lists

Pandoc has two types of lists: 'loose' (entries separted with newlines, each entry formatted as a paragraph) and 'soft' (no newlines between entries). Google docs has no distinction -- all lists are loose.


### Hyperlinks (and link hi-jacking)

Where HTML is the output format, all the links are hijacked.

```{r}
results %>% filter(test == "hyperlink.md")

read_txt("../test_cases/hyperlink.md")
read_txt("open_office_doc-ms_word_doc/returned-hyperlink.md")
read_txt("ms_word_doc-open_office_doc/returned-hyperlink.md")


cat(read_txt("../test_cases/list_bulleted.md"))
cat(read_txt("open_office_doc-ms_word_doc/returned-list_bulleted.md"))
cat(read_txt("ms_word_doc-open_office_doc/returned-list_bulleted.md"))
cat(read_txt("ms_word_doc-ms_word_doc/returned-list_bulleted.md"))

```




## Best performing format combinations

In terms of preserving the same AST, uploading as an open-office document, and downloading as an MS Word document produces the most faithful results overall across tests.

```{r}
results %>% group_by(upload_format, download_format) %>% 
  summarise(mu = mean(same_ast)) %>% 
  arrange(desc(mu))

results %>% group_by(upload_format, download_format) %>% 
  summarise(mu = mean(same_ast_elements)) %>% 
  arrange(desc(mu))
```

## Which features require which combinations?

Are there are any features which cannot be acheived with open office -> MS Word, but can be achieved with others?

Yes -- numbered lists are not preserved with any MS Word output, or any Rich Text input. However, they are preserved (down to the exact AST) with Open Office -> Open Office.

```{r}
ow_id <- results %>% 
  filter(
    upload_format == "open_office_doc" & download_format == "ms_word_doc"
  ) %>% select(combo_id) %>% unique() %>% .$combo_id

possible_features <- results %>% 
  filter(same_ast_elements) %>% select(test) %>% unique() %>% .[,1]

possible_features[!possible_features %in% 
  filter(results, same_ast_elements & combo_id == ow_id)$test]

results %>% filter(test == "list_numbered.md")

```


---------------------------

[^1]: EPUB is a promising format supported by both Pandoc and Google docs, which has not been considered here.
