# Run through

This file contains some code to try and use the googdown package in and end-to-end manner. It's not a formal test, but could inform them.


```{r}

unlink(".googdown/", recursive = TRUE, force = TRUE)
file.copy("original.Rmd", "test.Rmd", overwrite = TRUE)

library(googdown)
library(magrittr)

gd_auth()

cat(brocks::read_txt("test.Rmd"))

# ul <- gd_upload("test.Rmd")

ul <- gdoc_push("test.Rmd")

cat(brocks::read_txt("test.Rmd"))

# Make changes remotely

ul <- gd_pull("test.Rmd")

cat(brocks::read_txt("test.Rmd"))

unlink(".googdown/", recursive = TRUE, force = TRUE)

```

- Text which gets echo'ed completely goofs this up. You might want to have some kind of check for that (could consult knit_meta)

- You'll need to do some explicit input:output matching with figure captions. At the moment it just borks the whole thing.

- THere is something borking figure captions before they get sent to Google

- You need to do something about codeblocks which produce something
  - You can and should allow these to be deleted
  - You can and should have warnings if things change

- You could write something to fix up lists, if you wanted to





```{r}

# Obtain a file with pandoc codeblock
lines <- readLines(googdown:::ast_to_md("3-source.ast"))

pandoc_fenced_to_knitr_block(lines)

## TODO:
##  - Add spaces before and after headings, codeblocks, etc



# Lines which start with 3 backticks
backtick_lines <- (1:length(lines))[grepl("^```", lines)]

codeblock_starts <- backtick_lines[1:length(backtick_lines) %% 2 == 0]
codeblock_ends   <- backtick_lines[1:length(backtick_lines) %% 2 == 1]


brocks::read_txt("blah.md")

```


